### 2020-10-15 conclusion.tex
I messed with the related work and it became LONGER :'(

\begin{document}
\section{Related Work and Conclusion}
\textit{Simon: I think you really need some mention of MPSTs here, since a key motivation for MPSTs is ensuring deadlock-freedom between participants in a single MPST session. You can then talk about the global progress work by Coppo \etal, too, which incorporates an 'interaction typing system' (that few people understand). How does that relate?}

\subsubsection*{On deadlock freedom and progress in $\pi$-calculus and $\lambda$-calculus}

Deadlock freedom and progress are well-studied in the $\pi$-calculus. In the standard typed $\pi$-calculus, one line of work is Kobayashi's approach to deadlock-freedom~\cite{kobayashi98}, where priorities are abstract tags defined over a partially ordered set. These abstract tags were later simplified to natural numbers. Pairs of obligations and capabilities were used in the type system for lock freedom~\cite{kobayashi02,kobayashi06}, which allowed more $\pi$-calculus processes to be typed. Following, Kobayashi's line of work, Padovani~\cite{padovani13} adapted the obligation and capability pairs to session types, and later on he further simplified them to a single priority for linear $\pi$-calculus~\cite{padovani14}. By using the encoding of session types into linear types~\cite{kobayashi07,dardhagiachino12,dardha14beat,dardha16}, the priority-based technique for deadlock freedom can be transferred onto the $\pi$-calculus with session types.

The groundbreaking work on progress for session-typed $\pi$-calculus, by Dezani \etal~\cite{dezani-ciancaglinimostrous06}, guarantees progress by allowing only one active session at a time. In~\cite{dezani-ciancagliniliguoro09progress} the authors introduce a partial order on channels in line with Kobayashi's work~\cite{kobayashi98}. Carbone and Debois~\cite{carbonedebois10} define progress for session typed $\pi$-calculus in terms of a \emph{catalyser} used to provide a missing counterpart to a process, thus guaranteeing progress.
Carbone \etal~\cite{carbonedardha14} studied further the use of catalysers and showed that progress is a compositional form of lock-freedom. Following Padovani~\cite{padovani14}, the authors show that by using the encoding of session types~\cite{dardhagiachino12} and Kobayashi's obligations/capabilities, we can obtain progress for session types. Vieira and Vasconcelos~\cite{vieiravasconcelos13} used single priorities and an abstract partial order to guarantee deadlock freedom in a session-typed $\pi$-calculus.

Gay~\etal~\cite{gaynagarajan03} and Vasconcelos \etal~\cite{vasconcelosravara04,vasconcelosgay06} were the first to introduce a functional language with session types. However, such works and including the GV line~\cite{gayvasconcelos10,gayvasconcelos12} did not guarantee deadlock freedom. This property was later on addressed~\cite{lindleymorris15,wadler15} via syntactic restrictions allowing only communication in tree structures, as discussed in the introduction. Toninho \etal~\cite{toninhocaires12} present an encoding of simply-typed $\lambda$-calculus into session-typed $\pi$-calculus investigating the concurrency features of $\lambda$-calculus and providing a new logical explanation. Again, their focus is not on deadlock freedom.

\subsubsection*{On Curry-Howard correspondences}
The Curry-Howard correspondence between logic and types laid the foundation for functional programming~\cite{wadler15}. With the rise of linear logic~\cite{girard87} a Curry-Howard correspondence was established between linear logic and the linear $\pi$-calculus~\cite{abramsky94,bellinscott94}. With the rise of session types, a new correspondence between linear logic and the $\pi$-calculus was proposed first by Caires and Pfenning~\cite{cairespfenning10}, for dual intuitionistic linear logic (DILL), and later by Wadler~\cite{wadler15} for classical linear logic (CLL). The underlying language in both correspondences satisfy deadlock freedom by design, by forbidding cyclic connections via the logical cut rule. Dardha and Gay follow Padovani's work~\cite{padovani14} and use priorities in CP to define a new calculus which is deadlock-free by design, based on classical linear logic correspondence with session types~\cite{wadler12}. Dardha and P\'{e}rez~\cite{dardhaperez15} compare Kobayashi-style typing and CLL typing, and prove that CLL corresponds to Kobayashi's system with the restriction that only single cuts are allowed. Balzer \etal study sharing~\cite{balzerpfenning17} and later on deadlock freedom~\cite{balzertoninho19} in $\text{SILL}_S$ and $\text{SILL}_{S+}$, session-typed $\pi$-calculi based on the correspondence with DILL. In $\text{SILL}_S$~\cite{balzerpfenning17}, types are either linear or shared with modal operators connecting them, which allows sharing, but not deadlock freedom. In $\text{SILL}_{S+}$~\cite{balzertoninho19} types are enhanced with information regarding the order of acquiring resources, which guarantees deadlock freedom.

>>>>

\begin{document}
\section{Related Work and Conclusion}
\textit{Simon: I think you really need some mention of MPSTs here, since a key motivation for MPSTs is ensuring deadlock-freedom between participants in a single MPST session. You can then talk about the global progress work by Coppo~\etal, too, which incorporates an 'interaction typing system' (that few people understand). How does that relate?}

\subsubsection*{Deadlock freedom}

Deadlock freedom and progress are well-studied in the $\pi$-calculus.
For the standard typed $\pi$-calculus, one line of work is Kobayashi's approach to deadlock-freedom~\cite{kobayashi98}, where priorities are abstract tags defined over a partially ordered set. These abstract tags were later simplified to natural numbers. Pairs of obligations and capabilities were used in the type system for lock freedom~\cite{kobayashi02,kobayashi06}, which allowed more $\pi$-calculus processes to be typed. Padovani~\cite{padovani13} adapted the obligation and capability pairs to session types, and later simplified them to a single priority for linear $\pi$-calculus~\cite{padovani14}. Using the encoding of session types into linear types~\cite{kobayashi07,dardhagiachino12,dardha14beat,dardha16}, the priority-based technique for deadlock freedom can be transferred onto the $\pi$-calculus with session types.

For the session-typed $\pi$-calculus, the foundational work by Dezani~\etal~\cite{dezani-ciancaglinimostrous06} guarantees progress by allowing only one active session at a time. Dezani later~\cite{dezani-ciancagliniliguoro09progress} introduces a partial order on channels in line with Kobayashi's work~\cite{kobayashi98}. Carbone and Debois~\cite{carbonedebois10} define progress for session typed $\pi$-calculus in terms of a \emph{catalyser} used to provide a missing counterpart to a process, thus guaranteeing deadlock freedom.
Carbone~\etal~\cite{carbonedardha14} studied the use of catalysers and showed that progress is a compositional form of lock-freedom. Following Padovani~\cite{padovani14}, Dardha~\etal~\cite{dardhagiachino12} show that by using the encoding of session types and Kobayashi's obligations/capabilities, we can obtain progress for session types.\wen{I'm not sure what this means.} Vieira and Vasconcelos~\cite{vieiravasconcelos13} used single priorities and an abstract partial order to guarantee deadlock freedom in a session-typed $\pi$-calculus.

Gay~\etal~\cite{gaynagarajan03} and Vasconcelos~\etal~\cite{vasconcelosravara04,vasconcelosgay06} were the first to introduce a functional language with session types. However, such works, including early GV~\cite{gayvasconcelos10,gayvasconcelos12} did not guarantee deadlock freedom, until it was addressed via syntactic restrictions~\cite{lindleymorris15,wadler15}. Toninho~\etal~\cite{toninhocaires12} present a translation of simply-typed $\lambda$-calculus into session-typed $\pi$-calculus. However, their focus is not on deadlock freedom.

\subsubsection*{Ties with logic}
The correspondence between logic and types lays the foundation for functional programming~\cite{wadler15}. Since its inception by Girard~\cite{girard87}, linear logic has been a candidate for a foundational correspondence for concurrent programs. A~correspondence with linear $\pi$-calculus was established early on by Abramsky~\cite{abramsky94,bellinscott94}. A~correspondence between session-typed $\pi$-calculus and dual intuitionistic linear logic was developed by Caires and Pfenning~\cite[$\pi\text{DILL}$]{cairespfenning10}, and with classical linear logic by Wadler~\cite[CP]{wadler15}, both guaranteeing deadlock freedom as a result of their connection to logic. Dardha and Gay~\cite[PCP]{dardhagay18} integrate Padovani's work on priorities~\cite{padovani14} with CP, creating the first calculus which combines priorities and strong ties with logic. Dardha and P\'{e}rez~\cite{dardhaperez15} compare Kobayashi-style typing and CLL typing, and show that CLL corresponds to Kobayashi's system with the restriction that only single cuts are allowed.\wen{What do you mean by ``single cuts''?} Balzer~\etal~\cite[$\text{SILL}_S$]{balzerpfenning17} introduce shared state, which breaks deadlock freedom. They later restore deadlock freedom using priorities~\cite[$\text{SILL}_{S+}$]{balzertoninho19}.
